---
title: "model_building"
author: "Christian Schroeder"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(heatmaply)
library(faraway)
library(MASS)
library(tidyverse) 
```

Import the cleaned data

```{r}
melbourne <- read.csv("melbourne_cleaned.csv", header=TRUE)
melbourne$YearBuilt<-round(as.numeric(melbourne$YearBuilt), 0)

# ==============================================================================
# The dataset contains multiple categorical variables, so create a list of them
# and tell R to turn them into a factor
# ==============================================================================

cats <- c('Type', 'Regionname', 'Postcode', 'CouncilArea', 'Suburb')
melbourne[,cats] <- lapply(melbourne[,cats], factor)
attach(melbourne)
names(melbourne)
```

```{r}
full <- lm(Price~., data=melbourne)
summary(full)
```

-   drop Postcode, Method, Month, Latitude, Longitude, Regionname, Suburb, Date

```{r}
reduced <- lm(Price~Rooms+Type+Distance+Bathroom+Car+Landsize+BuildingArea+YearBuilt+Propertycount+CouncilArea+Year+SellerG)
summary(reduced)
```

-   Test if can drop the predictors

```{r}
anova(reduced,full)
```

-   We can drop Postcode, Method, Month, Latitude, Longitude, Regionname, Suburb, Date

Transform?

```{r}
# Check if transformation needed
bc <- boxcox(reduced, lambda = seq(-0.3, 0.3, 0.01))

# Because 1 is not within the range, the variance is not constant and it would be best to transform the response variable. Getting the recommended transformation value.
bc <- bc$x[which.max(bc$y)]
bc
```

-   transform

```{r}
# transform response and rebuild the model
Price.tran <- Price**bc
reduced.tran <-lm(Price.tran~Rooms+Type+Distance+Bathroom+Car+Landsize+BuildingArea+YearBuilt+Propertycount+CouncilArea+Year+SellerG)
summary(reduced.tran)
```

```{r}
# Check for linearity and constant variance
plot(reduced.tran$fitted.values,reduced.tran$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(reduced.tran$residuals, main="ACF of Residuals")

# Check normality
qqnorm(reduced.tran$residuals)
qqline(reduced.tran$residuals, col="red")

# Boxcox
boxcox(reduced.tran, lambda = seq(-5, 5, 0.01))
```

-   maybe normality is not met?

-   Should we now drop SellerG?

```{r}
reduced.tran2<-lm(Price.tran~Rooms+Type+Distance+Bathroom+Car+Landsize+BuildingArea+YearBuilt+Propertycount+CouncilArea+Postcode+Year)
anova(reduced.tran, reduced.tran2)
```

-   SellerG should be dropped

```{r}
summary(reduced.tran2)
```

-   Still need to transform a predictor?

-   Take the log of everything??

```{r}
reduced.tran3<-lm(Price.tran~Rooms+Type+Distance+Bathroom+Car+log(Landsize+0.01)+BuildingArea+YearBuilt+Propertycount+CouncilArea+Year)
summary(reduced.tran3)
```
-   Stick with tran2

-   Use cooks distance to view influential observations
```{r}
cooksd <- cooks.distance(reduced.tran2)
plot(cooksd, pch="*", cex=2, main="Influential Observations by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```
-   Check for multicollinearity
```{r}
library(faraway)
vif(reduced.tran2)
```

