---
title: "Forward Selection for Baseline Model"
author: "Will Tyree"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(heatmaply)
library(faraway)
library(MASS)
library(tidyverse) 
```

### Import Data, create full and intercept only models, perform forward selection

```{r}
melbourne <- read.csv("melbourne_cleaned.csv", header=TRUE)
melbourne = subset(melbourne, select = -c(X) )
cats <- c('Type', 'Regionname', 'Postcode', 'CouncilArea', 'Suburb', 'Year', 'Month')
melbourne[,cats] <- lapply(melbourne[,cats], factor)
melbourne$Population <- as.numeric(gsub(",","",melbourne$Population))

regnull <- lm(Price~1, data=melbourne)
regfull <- lm(Price~., data=melbourne)
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")
```

### Creating the baseline model from the forward selection results

```{r}
model <- lm(Price ~ BuildingArea + Suburb + Type + Landsize + 
    CouncilArea + Bathroom + YearBuilt + SellerG + Year + Rooms + 
    Method + Car + Walk.Score + Month + Lattitude + Distance, 
    data = melbourne)
summary(model)
```

### Checking regression assumptions

```{r}
plot(model$fitted.values,model$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(model$residuals, main="ACF of Residuals")

# Check normality
qqnorm(model$residuals)
qqline(model$residuals, col="red")
```

### Getting optimal box cox value to transform the response

```{r}
bc <- boxcox(model, lambda = seq(-0.3, 0.3, 0.01))

# Because 1 is not within the range, the variance is not constant and it would be best to transform the response variable. Getting the recommended transformation value.
bc <- bc$x[which.max(bc$y)]
bc
```

### Transforming the response with optimal BC

```{r}
Price.tran <- log(melbourne$Price)
model.tran <-lm(Price.tran~BuildingArea + Suburb + Type + Landsize + 
    CouncilArea + Bathroom + YearBuilt + SellerG + Year + Rooms + 
    Method + Car + Walk.Score + Month + Lattitude + Distance, 
    data = melbourne)
summary(model.tran)
```

### Checking assumptions on transformed model

```{r}
plot(model.tran$fitted.values,model.tran$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(model.tran$residuals, main="ACF of Residuals")

# Check normality
qqnorm(model.tran$residuals)
qqline(model.tran$residuals, col="red")
```

### Check for multicollinearity

```{r}
library(faraway)
vif(model.tran)
```

### Partial Regression Plots

```{r}
result.y.rooms <- lm(Price.tran~BuildingArea + Suburb + Type + Landsize + 
    CouncilArea + Bathroom + YearBuilt + SellerG + Year + 
    Method + Car + Walk.Score + Month + Lattitude + Distance, 
    data = melbourne)
result.rooms <- lm(Rooms~BuildingArea + Suburb + Type + Landsize + 
    CouncilArea + Bathroom + YearBuilt + SellerG + Year + 
    Method + Car + Walk.Score + Month + Lattitude + Distance, 
    data = melbourne)
res.y.rooms <- result.y.rooms$residuals
res.rooms <- result.rooms$residuals
plot(res.rooms, res.y.rooms, main="Partial Regression Plot of Rooms")
abline(h=0)
```

### Check normality for individual predictors that may need transformation

```{r}
par(mfrow=c(1,2))
qqnorm((melbourne$Rooms),main="Normal Q-Q Plot of Rooms");qqline((melbourne$Rooms))
qqnorm(log(melbourne$Rooms),main="Normal Q-Q Plot of log Rooms");qqline(log(melbourne$Rooms))
```

### Fitting model with rooms log transformed

```{r}
Price.tran <- log(melbourne$Price)
model.tran.room <-lm(Price.tran~BuildingArea + Suburb + Type + Landsize + 
    CouncilArea + Bathroom + YearBuilt + SellerG + Year + log(Rooms) + 
    Method + Car + Walk.Score + Month + Lattitude + Distance, 
    data = melbourne)
summary(model.tran.room)
```

### Checking assumptions again

```{r}
plot(model.tran.room$fitted.values,model.tran.room$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(model.tran.room$residuals, main="ACF of Residuals")

# Check normality
qqnorm(model.tran.room$residuals)
qqline(model.tran.room$residuals, col="red")
```

### Model Diagnostics

```{r}
res <- model.tran.room$residuals
student.res <- rstandard(model.tran.room)
ext.student.res <- rstudent(model.tran.room)

par(mfrow=c(1,3))
plot(model.tran.room$fitted.values, res, main="Residuals")
plot(model.tran.room$fitted.values, student.res, main="Studentized Residuals")
plot(model.tran.room$fitted.values, ext.student.res, main="Externally Studentized Residuals")
```

```{r}
n <- length(melbourne)
p <- 17
cv <- qt(1-0.05/(2*n), n-p-1)

# another way is to make a plot
plot(ext.student.res, main="Externally Studentized Residuals", ylim=c(-15,15))
abline(h=qt(1-0.05/(2*n), n-p-1), col = "red")
abline(h=-qt(1-0.05/(2*n), n-p-1), col = "red") # everything must be within line
ext.student.res[abs(ext.student.res)>qt(1-0.05/(2*n), n-p-1)] 
```

### Inspect outliers in the response - will need to find them in the raw data that contains the address

```{r}
melbourne_raw <- read.csv("melb_data_raw.csv", header=TRUE)
```

-   512 - correct
-   579 - 171 Moreland Rd, price listed as 145,000, actual price 1,550,000 <https://www.developmentready.com.au/properties/171-moreland-road-coburg-vic-3058>
-   1840 - correct
-   6458 - 30 Pyne St, listed as 131000, actual price 1350000

```{r}
melbourne$Price[579]=1550000
melbourne$Price[6458]=1350000
```

### Inspect leverage points

```{r}
lev<-lm.influence(model.tran.room)$hat
plot(lev, main="Leverages", ylim=c(0, 1.5))
abline(h=2*p/n, col="red")
```

### influential obervations

```{r}
COOKS <- cooks.distance(model.tran.room)
COOKS[COOKS>qf(0.5,p,n-p)]
```

## Hypothesis Testing: Adding Bathrooms

H0: Bathroom Coefficient = 11.01228 Ha: Bathroom Coefficient != 11.01228

```{r}
bathroomCOEF <- 0.06268
bathroomSE <- 0.003868

mean <- mean(melbourne$Price)
pctmean <- mean*0.057
logmeanprice <- log(pctmean)
t <- (bathroomCOEF-logmeanprice)/bathroomSE
abs(t)

critical <- qt(0.025, 12895)
critical
```
