---
title: "model_building"
author: "Christian Schroeder"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(heatmaply)
library(faraway)
library(MASS)
library(tidyverse) 
```

Import the cleaned data

```{r}
melbourne <- read.csv("melbourne_cleaned.csv", header=TRUE)
melbourne$YearBuilt<-round(as.numeric(melbourne$YearBuilt), 0)

# ==============================================================================
# The dataset contains multiple categorical variables, so create a list of them
# and tell R to turn them into a factor
# ==============================================================================

cats <- c('Type', 'Regionname', 'Postcode', 'CouncilArea', 'Suburb')
melbourne[,cats] <- lapply(melbourne[,cats], factor)
attach(melbourne)
names(melbourne)
```

# MLR

### Data Preparation

#### Location Predictor

-   We only need one location-based categorical predictor

    -   Suburb, Postcode, Regionname, or CouncilArea

```{r}
plot(Price~Regionname)
plot(Price~Suburb)
plot(Price~Postcode)
plot(Price~CouncilArea)
```

-   Focusing on granularity, Postcode and Suburb are the most specific location variables, but they would likely add too much complexity to the model
-   So we should choose between Regionname or CouncilArea

```{r}
region_classes <- melbourne %>% group_by(Regionname) %>% summarize(mean_price = round(mean(Price),2))
ggplot(region_classes, aes(x=Regionname, y=mean_price, fill=Regionname)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(label = mean_price), hjust = .5,
            color = "white",
            size = 3) + 
  theme_classic()

council_classes <- melbourne %>% group_by(CouncilArea) %>% summarize(mean_price = round(mean(Price),2))
ggplot(council_classes, aes(x=CouncilArea, y=mean_price, fill=CouncilArea)) +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(label = mean_price), hjust = .5,
            color = "white",
            size = 3) + 
  theme_classic()
```

-   Regions don't seem different enough in their mean prices to help a model estimate house price accurately
-   There is a lot more variability in mean prices based on council area, so we will use that.

#### Determining significant numeric predictors

-   Rooms, Distance, Bathroom, Car, Landsize, BuildingArea, YearBuilt, Month, Year

-   Look at relationships between Price and numeric predictors

```{r}
numericVals <- subset(melbourne, select=-c(SellerG, Method, X, Suburb, Date, Postcode, CouncilArea, Lattitude, Longtitude, Regionname, Type))

# Compute correlation coefficients
cor.coef <- cor(numericVals)
cor.coef
```

-   Price has the strongest correlation with BuildingArea, and Rooms coming in 2nd.
-   In order of correlation: BuildingArea, Rooms, Bathroom, YearBuilt, Landsize, Car, Distance, PropertyCount, Month, Year
-   Fit a model in that order

```{r}
model <- lm(Price~BuildingArea+Rooms+Bathroom+YearBuilt+Landsize+Car+Distance+Propertycount+Month+Year)
summary(model)
```

-   Propertycount is insignificant

-   53.61% of the variation in Price can be explained by the 9 predictors in the above model.

-   Which location-based categorical variable would explain more variability?

```{r}
reduced_model <- model
alt_model_1 <- lm(Price~Regionname)

alt_model_2 <- lm(Price~CouncilArea)

anova(reduced_model, alt_model_1)
anova(reduced_model, alt_model_2)
```

-   The partial f tests has the same P-values indicating that both predictors would contribute greatly to the model. Looking at the adjR2 of the each reduced model, we can assume that adding CouncilArea would increase the adjr2 a little more than Suburb

```{r}
full_model <- lm(Price~BuildingArea+Rooms+Bathroom+YearBuilt+Landsize+Car+Distance+Propertycount+Month+Year+CouncilArea)
summary(full_model)
```

-   Propertycount became significant after adding the councilarea
-   Check for autocorrelation

```{r}
pacf(full_model$residuals, main="PACF of Residuals")
```

-   There is autocorrelation at lag 1

-   Check for multicollinearity

```{r}
vif <- vif(full_model)
sort(vif,decreasing = T)
```

-   There is no multicollinearity in the full_model

```{r}
# Check for linearity and constant variance
plot(full_model$fitted.values,full_model$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(full_model$residuals, main="ACF of Residuals")

# Check normality
qqnorm(full_model$residuals)
qqline(full_model$residuals, col="red")
```

### How to increase AdjR2?
