---
title: "model_building"
author: "Christian Schroeder"
date: "5/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(heatmaply)
library(faraway)
library(MASS)
```

Import the cleaned data

```{r}
melbourne <- read.csv("melbourne_cleaned.csv", header=TRUE)
melbourne$YearBuilt<-round(as.numeric(melbourne$YearBuilt), 0)

# ==============================================================================
# The dataset contains multiple categorical variables, so create a list of them
# and tell R to turn them into a factor
# ==============================================================================

cats <- c('Type', 'Regionname', 'Postcode', 'CouncilArea', 'Suburb')
melbourne[,cats] <- lapply(melbourne[,cats], factor)
attach(melbourne)
names(melbourne)
```

Visualize the numeric correlations

```{r}
numericVals <- subset(melbourne, select=-c(SellerG, Method, X, Suburb, Date, Postcode, CouncilArea, Lattitude, Longtitude, Regionname, Type))

# Compute correlation coefficients
cor.coef <- cor(numericVals)
cor.coef

# Compute correlation p-values
cor.test.p <- function(x){
    FUN <- function(x, y) cor.test(x, y)[["p.value"]]
    z <- outer(
      colnames(x), 
      colnames(x), 
      Vectorize(function(i,j) FUN(x[,i], x[,j]))
    )
    dimnames(z) <- list(colnames(x), colnames(x))
    z
}

p <- cor.test.p(numericVals)
p
heatmaply_cor(
  cor.coef,
  node_type = "scatter",
  point_size_mat = -log(p), 
  point_size_name = "-log10(p-value)",
  label_names = c("x", "y", "Correlation")
)
```

### Fit SLR and Check Assumptions

First, we inspected this relationship by plotting Price against Rooms. Check the linear assumptions.

```{r}
slr <- lm(Price~Rooms)
summary(slr)

# Check for linearity and constant variance
plot(slr$fitted.values,slr$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(slr$residuals, main="ACF of Residuals")

# Check normality
qqnorm(slr$residuals)
qqline(slr$residuals, col="red")
```

1.  The mean value of y, is a straight-line function of x

    1.  The residuals are not evenly scattered across the horizontal axis, indicating a non-linear relationship between Price and Rooms.

2.  The errors have constant variance

    1.  The variance is not constant. There is a bulge in the variance of the residuals between 3 and 5 rooms, with an slight overall decrease as number of rooms goes up.

3.  The errors have mean 0

    1.  The errors do not have mean 0.

4.  The errors are independent

    1.  The ACF plot shows the errors at several levels have a correlation, so we cannot say the errors are independent.

5.  The errors have a normal distribution

    1.  The QQ plot indicates there is not normality of the residuals

### Transform the Response, Price

Because both the assumptions of linearity and constant variance are not met, we should look to stabilize the variance first by transforming the response variable.

-   The boxcox plot shows a 95% CI that does not contain 1, reinforcing that the variance is not constant and that we should transform the response variable.

-   Getting the optimal value to transform the response by, -0.130303

```{r}
# Check if transformation needed
bc <- boxcox(slr, lambda = seq(-0.3, 0, 0.01))

# Because 1 is not within the range, the variance is not constant and it would be best to transform the response variable. Getting the recommended transformation value.
bc <- bc$x[which.max(bc$y)]
bc
```

### Check Assumptions of Transformed Price Model

After transforming the predictor and refitting the model, we need to recheck all the regression assumptions.

```{r}
# transform response and rebuild the model
Price.tran <- Price**bc
slr.tran <-lm(Price.tran~Rooms)

# Check for linearity and constant variance
plot(slr.tran$fitted.values,slr.tran$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(slr.tran$residuals, main="ACF of Residuals")

# Check normality
qqnorm(slr.tran$residuals)
qqline(slr.tran$residuals, col="red")

# Boxcox
boxcox(slr.tran, lambda = seq(0.5, 1.5, 0.01))
```

1.  The mean value of y, is a straight-line function of x

    1.  The residuals are still not evenly scattered across the horizontal axis, indicating a non-linear relationship between transformed Price and Rooms.

2.  The errors have constant variance

    1.  The variance of the residuals is a lot more constant. 1 now lies within the 95% CI, so no further transformation to the response is needed.

3.  The errors have mean 0

    1.  There is still a noticeable pattern, so the errors do not have mean 0.

4.  The errors are independent

    1.  The ACF plot shows the errors at several levels still have a correlation, so we cannot say the errors are independent.

5.  The errors have a normal distribution

    1.  The QQ plot indicates there is normality of the residuals

### Transform the Predictor, BuildingArea

Because the residuals are not evenly scattered across the horizontal axis, the linearity assumption is not met and we need to transform the predictor.

-   Do a log transformation of BuildingArea

```{r}
# transform predictor and rebuild the model
Rooms.tran <- log(Rooms)
slr.tran2 <-lm(Price.tran~Rooms.tran)

# Check for linearity and constant variance
plot(slr.tran2$fitted.values,slr.tran2$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(slr.tran2$residuals, main="ACF of Residuals")

# Check normality
qqnorm(slr.tran2$residuals)
qqline(slr.tran2$residuals, col="red")

# Boxcox
boxcox(slr.tran2)
```

Transform the response again?

```{r}
bc2 <- boxcox(slr.tran2)
bc2 <- bc2$x[which.max(bc2$y)]
bc2
```

```{r}
# transform response and rebuild the model
Price.tran2 <- Price.tran**bc2
slr.tran3 <-lm(Price.tran2~Rooms.tran)

# Check for linearity and constant variance
plot(slr.tran3$fitted.values,slr.tran3$residuals, main="Plot of Residuals against Fitted Values")
abline(h=0,col="red")

# Check for independence
acf(slr.tran3$residuals, main="ACF of Residuals")

# Check normality
qqnorm(slr.tran3$residuals)
qqline(slr.tran3$residuals, col="red")

# Boxcox
boxcox(slr.tran3, lambda = seq(0.5, 1.5, 0.01))
```

-   All assumptions met except for independent errors, ACF plot show significant lag
